Ages of Stars and Planets in the Kepler Field Over the First Four Billion Years

### 2024-08-30 papers/paper/ms.tex
update ms

% upon AAS submission
%\documentclass[11pt,twocolumn,tighten,linenumbers]{aastex63}
%\documentclass[11pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
% drafting / arxiv
\documentclass[11pt,twocolumn,tighten]{aastex63}
\turnoffedit

@@

% <250 words
\begin{abstract}
  %
  Recent analyses of FGK stars in open clusters have helped clarify
  the precision with which a star's rotation rate and lithium content
  can be used as empirical indicators for its age.
  %
  Here we apply this knowledge to stars observed by Kepler.
  %
  Rotation periods are drawn from previous work; lithium is measured
  from new and archival Keck/HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  (harboring \nplwgyroagenograzing\ known planets) for which our method is
  applicable.
  %
  We find that our rotational ages accurately recover the ages of
  stars in open clusters spanning 0.04-2.5\,Gyr; they also agree with
  $\gtrsim$90\% of the independent lithium ages.
  %
  The resulting yield includes \nplyounggyrotwosigma\ planets younger
  than 1\,Gyr at 2$\sigma$, and \nplyounggyro\ with median ages below
  1\,Gyr.
  %
  This is about half the number expected under the classic assumption
  of a uniform star formation history.
  %
  \deleted{We find that the scarcity of sub-gigayear systems can be
  attributed to the star formation rate in the Kepler field dropping
  by a factor of \ratiosfr$\pm$\uncratiosfr over the past 3\,Gyr.  We
  observe this trend both for known planet hosts and for the parent
  stellar sample.  This ``demographic cliff'' in the Galaxy's star
  formation history has been previously reported, and its confirmation
  helps clarify the age distribution of the known transiting
  exoplanets.}
  %
  \added{The age distribution that we observe, rather than being uniform,
  shows that the youngest stars in the Kepler field are 3-5 times
  rarer than stars 3\,Gyr old.
  This trend holds for both known planet hosts and for the parent
  stellar sample.
  We attribute the existence of this ``demographic cliff'' to a
  combination of kinematic heating and a declining star formation rate
  in the Galaxy's thin disk, and highlight its impact on the age
  distribution of known transiting exoplanets.
  }
\end{abstract}

@@

\begin{enumerate}[label={\it \roman*)},leftmargin=12pt,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item The Kepler Object of Interest catalog, and our vetting of
    false positives within it, has now reached maturity
    \citep[e.g.][]{Thompson_2018}.
  %
  \item Measured rotation periods of FGK stars in open clusters now
    show not only the average evolution of $P_{\rm rot}(T_{\rm
    eff},t)$, but also how the astrophysical dispersion of stars
    around this average converges by the $\approx$700\,Myr age of
    Praesepe
    \citep[e.g.][]{Curtis_2019_ngc6811,Gillen_2020,Rampalli_2021,Fritzewski_2021,Rebull_2022,Dungee_2022,2023AJ....166...14B}.
  %
  \item Using open cluster data, we can marginalize over the range of
    ages that might explain any one star's rotation period
    \citep{Bouma_2023}.  This represents an improvement in the
    accuracy of uncertainty propagation relative to previous
    calibrations.
  %
  \item We can identify \replaced{binary stars}{stars that are
    binaries and that have high metallicity} with greater fidelity\deleted{ than
    in the past}, which can clarify otherwise problematic estimates of
    rotation-based ages.
  %
  \item New open clusters in various stages of dissolution have been
    found in the Kepler field
    \cite[e.g.][]{2019AJ....158..122K,Kerr_2021,Barber_2022}.
    These clusters offer stars that we can use to test
    the reliability of the available single-star age-dating methods.
\end{enumerate}

@@

We discuss our method for selecting the star and planet samples in
Section~\ref{sec:selection}, and describe the origin of our adopted
stellar parameters other than ages in Section~\ref{sec:stellarprops}.
We describe our age-dating methods in Sections~\ref{sec:rotage}
and~\ref{sec:liage}, and test them in
Section~\ref{sec:litagecomparison} using clusters in the Kepler field.
We discuss population-level trends in Sections~\ref{sec:results}
and~\ref{sec:disc}, and offer a few conclusions in
Section~\ref{sec:conclusions}.

@@

	\vspace{-0.35cm}
    \subfloat{
        \includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_preciseagesonly.pdf}
    }
%
%	\vspace{-0.35cm}
%	\subfloat{
%		\includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_dropfracshortrot.pdf}
%	}
  \end{center}
  \vspace{-0.6cm}
  \caption{{\bf Kepler's demographic cliff,} visible in the
  rotation-derived age distributions of its stars (left) and planet
  hosts (middle).  The top row shows \replaced{all stars with temperatures of
  3800--6200\,K for which we calculated rotational ages}{stars with temperatures of 3800--6200\,K}.  Opaque lines
  impose quality cuts on binarity, \added{metallicity, }crowding, and the star's
  evolutionary state; transparent lines do not\added{ (see
  Section~\ref{subsec:flags})}.  The bottom row shows
  stars with temperatures of 4400--5400\,K, which \replaced{have more precise
  ages due to their fast spin-down}{usually have more precise ages}.  The statistical uncertainty for a
  mean star at 1, 2, and 3\,Gyr is shown, and is identical across each
  row.  
  Finally, the right panel compares the rotation-derived ages
  against star formation histories derived using CMD fitting (dashed
  and dotted lines).  Completeness in Kepler's $P$$_{\rm rot}$
  detection sensitivity is near unity until $t$$\lesssim$3\,Gyr\added{, and decreases rapidly at older ages}
  \citep{2022ApJ...937...94M}.
  \label{fig:hist_tgyro}
  }
\end{figure*}

@@

{\it Cep-Her}---The Cep-Her complex ($\approx$40\,Myr) contains
Kepler-1627Ab, Kepler-1643b, Kepler-1974b and Kepler-1975\,Ab
\citep{Bouma_2022a,Bouma_2022b}.  Different sub-groups of the complex
vary in age by $\approx$50\% (Kerr~et al.,~submitted).  The planets
themselves are all on close-in orbits (5-25\,days), with sizes from
2-4\,$R_\oplus$.  Our rotation and lithium ages agree with the cluster
age for Kepler-1627A, Kepler-1974, and Kepler-1975A.  For Kepler-1643,
the rotation and cluster ages agree, and the lithium age is
2.0$\sigma$ above the cluster age reported by \citet{Bouma_2022b} for
RSG-5.

@@

\begin{figure}[!t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_anyyoung_anyyoung_allowsgrazingandhighruwe.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf Sizes, orbital periods, and rotation-based ages of transiting
    exoplanets younger than one billion years.}
    \replaced{Star symbols denote \nplyounggyrotwosigmanograzingnoruwe\ planets
    that are characterized in this work to have $t_{\rm gyro}$<1\,Gyr
    at $2$$\sigma$; circles are planets from the literature meeting
    the same age requirement, but with a heterogeneous set of age
    provenances.  Gray points are transiting planets from the NASA
    Exoplanet Archive older than 1\,Gyr.  If one were to instead
    select for precise ages ($t/\sigma_{\rm t}>3$), this would hide
    $\approx$ten $<$0.5\,Gyr planets, and add $\approx$fifteen
    0.5-1\,Gyr planets.  For this plot, we require our Kepler stellar
    hosts to not have $Q_{\rm star}$ bits 0--9 raised,
    and for the planets to similarly no have quality flags
    raised.}{Transiting systems are selected to have $t$<1\,Gyr at $2$$\sigma$.
    Star symbols denote \nplyounggyrotwosigmanograzingnoruwe\ planets
    characterized in this work to have such ages at the highest
    confidence, meaning that none of $Q_{\rm star}$ bits 0--10 are raised,
    and the planets have no quality flags
    raised.  Transparent crosses denote \nplyounggrazingorhighruwe\
    planets for which either the planet is grazing, or the star has
    RUWE$>$1.4.  Gray points are from the NASA Exoplanet Archive. }
    \label{fig:rp_period_age_results}
  }
\end{figure}

@@

Figure~\ref{fig:rp_period_age_results} shows the planets and planet
candidates with $t_{\rm gyro}$$<$1\,Gyr at 2$\sigma$.  \replaced{This}{The stars in this} figure
include\deleted{s} \nconfirmedplyounggyrotwosigmanograzingnoruwe\ confirmed
Kepler planets, and \ncandidateplyounggyrotwosigmanograzingnoruwe\
candidate planets, all of which are listed in Table~\ref{tab:planets}.
These include \nearthshighq\ objects Earth-sized or smaller,
\njupitershighq\ Jovian-sized planet candidates, \nsubsaturnshighq\
super-Neptunes (4--10\,$R_\oplus$), \nminineptuneshighq\ mini-Neptunes
above the radius valley as parametrized by
\citealt{2018MNRAS.479.4786V}, and \nsuperearthshighq\ super-Earth
sized planets below the radius valley.  While most of these planets
are on orbital periods below 50\,days, \nlongperiodhighq\ are on more
distant orbits.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=.99\textwidth]{threepanel_trilegal_comparison_twostepsfr.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{{\bf Impact of star formation rate (SFR) and kinematic heating on age distribution.}   
		Most Kepler stars are a few hundred parsecs above the Galactic plane ($\hat{Z}$ points to the north galactic pole;  $\hat{Y}$ is the direction of
		galactic rotation).  This is comparable to the $\approx$300\,pc mass-weighted scale height.
		The dotted gray and solid orange lines show synthetic age distributions
		from TRILEGAL assuming a constant SFR (middle) and a two-step
		SFR (right), described in the text.
		A combination of both kinematic heating and
		a decreasing SFR qualitatively match the observed paucity of the youngest stars.
		\label{fig:trilegal}
	}
\end{figure*}

@@

\added{Two separate astrophysical effects might explain our observed 
age distribution (Figure~\ref{fig:hist_tgyro}): a declining star formation
rate (SFR), and kinematic heating.   Kinematic heating is associated with an
increase in stellar velocity dispersions and the disk's vertical 
scale height over time \citep[e.g.][]{Villumsen1983,Schmidt2024}.
The Kepler field points toward $(l,b)$=(76.3$^\circ$,13.5$^\circ$).
A typical Kepler star at 1\,kpc is therefore $\approx$230\,pc
above the galactic plane, a significant fraction of the thin disk's 
mass-weighted $\approx$300\,pc scale height \citep{Bland-Hawthorn2016}.
}

\added{We assessed the relative importance of these two effects
	using the TRILEGAL Monte Carlo population synthesis tool }\citep[v1.6;][]{Girardi2005}.
\added{TRILEGAL treats the Galaxy as a linear combination of a thin disk, thick disk,
	halo, and bulge, and allows for two possible SFRs: constant, and a ``two-step''
	SFR in which the thin disk SFR is 50\% larger from 1-4\,Gyr than at other ages.
	This two-step SFR was introduced to improve the model's ability to
	match star counts from Hipparcos and 2MASS }\citep[see Section~5.1 of][]{Girardi2005}.
\added{For kinematic heating, TRILEGAL assumes that
	the disk scale height grows as
$h_{\rm z} = z_0 (1 + t/t_0)^\alpha$, for $\alpha$=5/3, $z_0$=94.7\,pc,
	and $t_0$=5.55\,Gyr, similar to \citet{Rana1991}.
	We queried the web interface toward the
	Kepler field, and toward the galactic plane (at $l$=76.3$^\circ$) to
	negate the effects of kinematic heating.
	We limited the output to stars with apparent Kepler magnitudes brighter than 16,
	and set the field of view to 10 ${\rm deg}^2$, the maximum
	possible through the web interface.
%	and set the Sun at $z_\odot$=20.8\,pc, $r$=8122\,pc from the Galactic center.
}

\added{Figure~\ref{fig:trilegal} compares the simulation results against
	our observed age distribution.
	The middle panel shows that assuming a constant SFR,
	kinematic heating alone can yield a two-fold difference
	in the abundance of the youngest stars relative to those $\approx$3\,Gyr old.
	The right-most panel
	shows that additionally including 
	the time-variable SFR yields the best agreement with the slope of
	our observed age distribution.
%	An important note is that the previously noted parameters
%	that were fine-tuned to fit different datasets, primarily photometric star counts
%	along various sight-lines.  The exercise appears to have been predictive.
}

@@

\deleted{Classifying Kepler stars as thin vs.\ thick disk members helps connect
them to previous work on the Galaxy's star formation history.  On
extragalactic scales, the star formation rate in spiral galaxies with
mass similar to the Milky Way peaked $\approx$10\,Gyr ago, and has
since decreased by an order of magnitude
\citep[e.g.][]{2004Natur.428..625H,2006ApJ...651..142H}.}
\added{Despite the expected differences between pencil-beam surveys
and volume-complete samples, our observed age distribution qualitatively
agrees with a number of previous studies.}
\replaced{In our own
galaxy, previous studies have}{Previous work has} measured star formation histories
through CMD fitting of resolved stellar populations
\citep[][]{2019A&A...624L...1M,2020NatAs...4..965R,2021MNRAS.501..302A,2022Natur.603..599X},
and by modeling the local white dwarf luminosity function
\citep[e.g.][]{2019ApJ...878L..11I}.  \citet{2019A&A...624L...1M} and
\citet{2019ApJ...878L..11I} for instance focused on stars (and white
dwarfs) within a few hundred parsecs, and both found a peak in the
local star formation rate 2-3\,Gyr ago.  \citet{2020NatAs...4..965R}
focused on stars in a wider 2\,kpc bubble, and additionally reported
three local maxima in the SFR, which they
associated with pericenter passages of the Sagittarius satellite
galaxy.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_dropfracshortrot.pdf}
		}
	\end{center}
	\vspace{-0.6cm}
	\caption{{\bf Potential impact of unresolved binaries.}   This figure shows
		the same raw (thin blue line) and quality-flag cleaned (``gyro applicable'') distributions
		as in Figure~\ref{fig:hist_tgyro}.
		However, even after applying our quality cuts, up to
		\fivepctnuniqstarsantosrotgyroappl\ out of \ngyroappllttendays\ ``clean'' rotators with
		$P_{\rm rot}$$<$10\,days could be unresolved $\lesssim$0.5\,AU  binaries (see Section~\ref{subsec:caveats}).
		The thick blue ``5\% binary contamination'' distribution corrects for this possibility by randomly removing
		what we estimate (estm.) to be the maximum possible number of such systems; the true
		correction might be smaller.  We expect such a correction to be more important in
		the sample of all Kepler stars than in the KOI host sample.
		\label{fig:hist_tgyro_corr}
	}
\end{figure*}

@@

{\it Searches for new associations}:
Some of the youngest stars in Tables~\ref{tab:planets}
and~\ref{tab:stars} have been linked to their birth clusters.
Others have not. A systematic search for the birth clusters of the
youngest ``field stars'' could combine positions and velocities from
Gaia with rotation measurements from TESS. Given the $\sim$100\,Myr
decoherence time of young clusters, the youngest ``field stars'' may
have easily identifiable young neighbors.

@@

Tables~\ref{tab:planets} and~\ref{tab:stars} summarize the results for
planets and stars, respectively.  Our recovered ages are accurate for
all 14 known Kepler planets in clusters.  While lithium provided
minimal added information for most of the sample\added{ due to its astrophysical
depletion rate}, $t_{\rm Li}$ and
$t_{\rm gyro}$ agreed in $\gtrsim$90\% of cases for which comparison
was possible.  Our \replaced{resulting ages}{results} included two-sided $t_{\rm gyro}$
and $t_{\rm Li}$ for \ltonegyrhighqconfirmedtwosided\ sub-gigayear
confirmed planets, and two-sided $t_{\rm gyro}$ with one-sided $t_{\rm
Li}$ for \ltonegyrhighqconfirmedonesided\ sub-gigayear confirmed
planets.  Allowing for ``candidate'' planets, grazing transit
geometries, and stars with RUWE far from unity expands the counts by a
factor of two.

>>>>

% upon AAS submission
%\documentclass[11pt,twocolumn,tighten,linenumbers]{aastex63}
\documentclass[11pt,twocolumn,tighten,linenumbers,trackchanges]{aastex63}
% drafting / arxiv
%\documentclass[11pt,twocolumn,tighten]{aastex63}
\turnoffedit

@@

% <250 words
\begin{abstract}
  %
  Recent analyses of FGK stars in open clusters have helped clarify
  the precision with which a star's rotation rate and lithium content
  can be used as empirical indicators for its age.
  %
  Here we apply this knowledge to stars observed by Kepler.
  %
  Rotation periods are drawn from previous work; lithium is measured
  from new and archival Keck/HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  (harboring \nplwgyroagenograzing\ known planets) for which our method is
  applicable.
  %
  We find that our rotational ages \deleted{accurately }recover the ages of
  stars in open clusters spanning 0.04-2.5\,Gyr; they also agree with
  $\gtrsim$90\% of the independent lithium ages.
  %
  The resulting yield includes \nplyounggyrotwosigma\ planets younger
  than 1\,Gyr at 2$\sigma$, and \nplyounggyro\ with median ages below
  1\,Gyr.
  %
  This is about half the number expected under the classic assumption
  of a uniform star formation history.
  %
  \deleted{We find that the scarcity of sub-gigayear systems can be
  attributed to the star formation rate in the Kepler field dropping
  by a factor of \ratiosfr$\pm$\uncratiosfr over the past 3\,Gyr.  We
  observe this trend both for known planet hosts and for the parent
  stellar sample.  This ``demographic cliff'' in the Galaxy's star
  formation history has been previously reported, and its confirmation
  helps clarify the age distribution of the known transiting
  exoplanets.}
  %
  \added{The age distribution that we observe, rather than being uniform,
  shows that the youngest stars in the Kepler field are 3-5 times
  rarer than stars 3\,Gyr old.
  This trend holds for both known planet hosts and for the parent
  stellar sample.
  We attribute the existence of this ``demographic cliff'' to a
  combination of kinematic heating and a declining star formation rate
  in the Galaxy's thin disk, and highlight its impact on the age
  distribution of known transiting exoplanets.
  }
\end{abstract}

@@

\begin{enumerate}[label={\it \roman*)},leftmargin=12pt,topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
  \item The Kepler Object of Interest catalog, and our vetting of
    false positives within it, has now reached maturity
    \citep[e.g.][]{Thompson_2018}.
  %
  \item Measured rotation periods of FGK stars in open clusters now
    show not only the average evolution of $P_{\rm rot}(T_{\rm
    eff},t)$, but also how the astrophysical dispersion of stars
    around this average converges by the $\approx$700\,Myr age of
    Praesepe
    \citep[e.g.][]{Curtis_2019_ngc6811,Gillen_2020,Rampalli_2021,Fritzewski_2021,Rebull_2022,Dungee_2022,2023AJ....166...14B}.
  %
  \item Using open cluster data, we can marginalize over the range of
    ages that might explain any one star's rotation period
    \citep{Bouma_2023}.  This represents an improvement in the
    accuracy of uncertainty propagation relative to previous
    calibrations.
  %
  \item We can identify \replaced{binary stars}{stars that are
    binaries and that have high metallicity} with greater fidelity\deleted{ than
    in the past}, which can clarify otherwise problematic estimates of
    rotation-based ages.
  %
  \item New open clusters in various stages of dissolution have been
    found in the Kepler field
    \cite[e.g.][]{2019AJ....158..122K,Kerr_2021,Barber_2022}.
    These clusters offer stars that we can use to test
    the reliability of \replaced{the available single-star}{our} age-dating methods.
\end{enumerate}

@@

We discuss our method for selecting the star and planet samples in
Section~\ref{sec:selection}, and describe the origin of our adopted
stellar parameters other than ages in Section~\ref{sec:stellarprops}.
We present our age-dating methods in Sections~\ref{sec:rotage}
and~\ref{sec:liage}, and test them in
Section~\ref{sec:litagecomparison} using clusters in the Kepler field.
We discuss population-level trends in Sections~\ref{sec:results}
and~\ref{sec:disc}, and offer a few conclusions in
Section~\ref{sec:conclusions}.

@@

	\vspace{-0.35cm}
    \subfloat{
        \includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_preciseagesonly.pdf}
    }
%
%	\vspace{-0.35cm}
%	\subfloat{
%		\includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_dropfracshortrot.pdf}
%	}
  \end{center}
  \vspace{-0.6cm}
  \caption{{\bf Kepler's demographic cliff,} visible in the
  rotation-derived age distributions of its stars (left) and planet
  hosts (middle).  The top row shows \replaced{all stars with temperatures of
  3800--6200\,K for which we calculated rotational ages}{stars with temperatures of 3800--6200\,K}.  Opaque lines
  impose quality cuts on binarity, \added{metallicity, }crowding, and the star's
  evolutionary state; transparent lines do not\added{ (see
  Section~\ref{subsec:flags})}.  The bottom row shows
  stars with temperatures of 4400--5400\,K, which \replaced{have more precise
  ages due to their fast spin-down}{have more precise ages}.  The statistical uncertainty for a
  mean star at 1, 2, and 3\,Gyr is shown, and is identical across each
  row.  
  Finally, the right panel compares the rotation-derived ages
  against star formation histories derived using CMD fitting (dashed
  and dotted lines).  Completeness in Kepler's $P$$_{\rm rot}$
  detection sensitivity is near unity until $t$$\lesssim$3\,Gyr\added{, and decreases rapidly at older ages}
  \citep{2022ApJ...937...94M}.
  \label{fig:hist_tgyro}
  }
\end{figure*}

@@

{\it Cep-Her}---The Cep-Her complex ($\approx$40\,Myr) contains
Kepler-1627Ab, Kepler-1643b, Kepler-1974b and Kepler-1975\,Ab
\citep{Bouma_2022a,Bouma_2022b}.  Different sub-groups of the complex
vary in age by $\approx$50\% \citep{Kerr2024}.  The planets
themselves are all on close-in orbits (5-25\,days), with sizes from
2-4\,$R_\oplus$.  Our rotation and lithium ages agree with the cluster
age for Kepler-1627A, Kepler-1974, and Kepler-1975A.  For Kepler-1643,
the rotation and cluster ages agree, and the lithium age is
2.0$\sigma$ above the cluster age reported by \citet{Bouma_2022b} for
RSG-5.

@@

\begin{figure}[!t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_anyyoung_anyyoung_allowsgrazingandhighruwe.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf Sizes, orbital periods, and rotation-based ages of transiting
    exoplanets younger than one billion years.}
    \replaced{Star symbols denote \nplyounggyrotwosigmanograzingnoruwe\ planets
    that are characterized in this work to have $t_{\rm gyro}$<1\,Gyr
    at $2$$\sigma$; circles are planets from the literature meeting
    the same age requirement, but with a heterogeneous set of age
    provenances.  Gray points are transiting planets from the NASA
    Exoplanet Archive older than 1\,Gyr.  If one were to instead
    select for precise ages ($t/\sigma_{\rm t}>3$), this would hide
    $\approx$ten $<$0.5\,Gyr planets, and add $\approx$fifteen
    0.5-1\,Gyr planets.  For this plot, we require our Kepler stellar
    hosts to not have $Q_{\rm star}$ bits 0--9 raised,
    and for the planets to similarly no have quality flags
    raised.}{Systems are selected to have $t$<1\,Gyr at $2$$\sigma$.
    Star symbols denote \nplyounggyrotwosigmanograzingnoruwe\ planets
    characterized in this work to have such ages at the highest
    confidence, meaning that none of $Q_{\rm star}$ bits 0--10 are raised,
    and the planets have no quality flags
    raised.  Transparent crosses denote \nplyounggrazingorhighruwe\
    planets for which either the planet is grazing, or the star has
    RUWE$>$1.4.  Gray points are from the NASA Exoplanet Archive. }
    \label{fig:rp_period_age_results}
  }
\end{figure}

@@

Figure~\ref{fig:rp_period_age_results} shows the planets and planet
candidates with $t_{\rm gyro}$$<$1\,Gyr at 2$\sigma$.  \replaced{This}{The stars in this} figure
include\deleted{s} \nconfirmedplyounggyrotwosigmanograzingnoruwe\ confirmed
Kepler planets, and \ncandidateplyounggyrotwosigmanograzingnoruwe\
candidate planets, all of which are listed in Table~\ref{tab:planets}.
These include \nearthshighq\ objects Earth-sized or smaller,
\njupitershighq\ Jovian-sized planet candidates, \nsubsaturnshighq\
super-Neptunes (4--10\,$R_\oplus$), \nminineptuneshighq\ mini-Neptunes
above the radius valley as parametrized by
\citealt{2018MNRAS.479.4786V}, and \nsuperearthshighq\ super-Earth
sized planets below the radius valley.  While most of these planets
have orbital periods below 50\,days, \nlongperiodhighq\ are on more
distant orbits.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=.99\textwidth]{threepanel_trilegal_comparison_twostepsfr.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{{\bf Impact of star formation rate (SFR) and kinematic heating on age distribution.}   
		Most Kepler stars are a few hundred parsecs above the Galactic plane ($\hat{Z}$ points to the north galactic pole;  $\hat{Y}$ toward the direction of
		galactic rotation).  This is comparable to the $\approx$300\,pc mass-weighted scale height.
		The dotted gray and solid orange lines show synthetic age distributions
		from TRILEGAL assuming a constant SFR (middle) and a two-step
		SFR (right), described in the text.
		A combination of both kinematic heating and
		a decreasing SFR are needed to qualitatively match the observed paucity of the youngest stars.
		\label{fig:trilegal}
	}
\end{figure*}

@@

\added{Two separate astrophysical effects might explain our observed 
age distribution (Figure~\ref{fig:hist_tgyro}): a declining star formation
rate (SFR), and kinematic heating.   Kinematic heating is associated with an
increase in both stellar velocity dispersions and the disk's vertical 
scale height over time \citep[e.g.][]{Villumsen1983,Schmidt2024}.
The Kepler field points toward $(l,b)$=(76.3$^\circ$,13.5$^\circ$).
A typical Kepler star at 1\,kpc is therefore $\approx$230\,pc
above the galactic plane, a significant fraction of the thin disk's 
mass-weighted $\approx$300\,pc scale height \citep{Bland-Hawthorn2016}.
}

\added{We assessed the relative importance of these two effects
	using the TRILEGAL Monte Carlo population synthesis tool }\citep[v1.6;][]{Girardi2005}.
\added{TRILEGAL treats the Galaxy as a linear combination of a thin disk, thick disk,
	halo, and bulge, and allows for two possible SFRs: constant, and a ``two-step''
	SFR in which the thin disk SFR is 50\% larger from 1-4\,Gyr than at other ages.
	This two-step SFR was introduced to improve the model's ability to
	match star counts from Hipparcos and 2MASS }\citep[see Section~5.1 of][]{Girardi2005}.
\added{For kinematic heating, TRILEGAL assumes that
	the disk scale height grows as
$h_{\rm z} = z_0 (1 + t/t_0)^\alpha$, for $\alpha$=5/3, $z_0$=94.7\,pc,
	and $t_0$=5.55\,Gyr, similar to \citet{Rana1991}.
	We queried the web interface toward the
	Kepler field, and toward the galactic plane (at $l$=76.3$^\circ$) to
	negate the effects of kinematic heating.
	We limited the output to stars with apparent Kepler magnitudes brighter than 16,
	and set the field of view to 10\,${\rm deg}^2$, the maximum
	available through the web interface.
%	and set the Sun at $z_\odot$=20.8\,pc, $r$=8122\,pc from the Galactic center.
}

\added{Figure~\ref{fig:trilegal} compares the results against
	our observed age distribution.
	The middle panel shows that assuming a constant SFR,
	kinematic heating alone can yield a two-fold difference
	in the abundance of the youngest stars relative to those $\approx$3\,Gyr old.
	The right-most panel
	shows that additionally including 
	the time-variable SFR yields the best agreement with the slope of
	our observed age distribution.
%	An important note is that the previously noted parameters
%	that were fine-tuned to fit different datasets, primarily photometric star counts
%	along various sight-lines.  The exercise appears to have been predictive.
}

@@

\deleted{Classifying Kepler stars as thin vs.\ thick disk members helps connect
them to previous work on the Galaxy's star formation history.  On
extragalactic scales, the star formation rate in spiral galaxies with
mass similar to the Milky Way peaked $\approx$10\,Gyr ago, and has
since decreased by an order of magnitude. }
%\citep[e.g.][]{2004Natur.428..625H,2006ApJ...651..142H}.}
\added{Despite the expected differences between pencil-beam surveys
and volume-complete samples, our observed age distribution qualitatively
agrees with a number of previous studies.}
\replaced{In our own
galaxy, previous studies have}{Previous work has} measured star formation histories
through CMD fitting of resolved stellar populations
\citep[][]{2019A&A...624L...1M,2020NatAs...4..965R,2021MNRAS.501..302A,2022Natur.603..599X},
and by modeling the local white dwarf luminosity function
\citep[e.g.][]{2019ApJ...878L..11I}.  \citet{2019A&A...624L...1M} and
\citet{2019ApJ...878L..11I} for instance focused on stars (and white
dwarfs) within a few hundred parsecs, and both found a peak in the
local star formation rate 2-3\,Gyr ago.  \citet{2020NatAs...4..965R}
focused on stars in a wider 2\,kpc bubble, and additionally reported
three local maxima in the SFR, which they
associated with pericenter passages of the Sagittarius satellite
galaxy.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=.98\textwidth]{comp_hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240821_maxage4000_dropfracshortrot.pdf}
		}
	\end{center}
	\vspace{-0.6cm}
	\caption{{\bf Potential impact of unresolved binaries.}   This figure shows
		the same raw (thin blue line) and quality-flag cleaned (``gyro applicable'') distributions
		as in Figure~\ref{fig:hist_tgyro}.
		However, even after applying our quality cuts, up to
		\fivepctnuniqstarsantosrotgyroappl\ of \ngyroappllttendays\ ``clean'' rotators with
		$P_{\rm rot}$$<$10\,days could be unresolved $\lesssim$0.5\,AU  binaries (see Section~\ref{subsec:caveats}).
		The thick blue ``5\% binary contamination'' distribution corrects for this possibility by randomly removing
		what we estimate (estm.) to be the maximum possible number of such systems; the true
		correction might be smaller.  We expect such a correction to be more important in
		the sample of all Kepler stars than in the KOI host sample.
		\label{fig:hist_tgyro_corr}
	}
\end{figure*}

@@

{\it Searches for new associations}:
Some of the youngest stars in Tables~\ref{tab:planets}
and~\ref{tab:stars} have been linked to their birth clusters.
Others have not. A systematic search for the birth clusters of the
youngest ``field stars'' could combine positions and velocities from
Gaia with rotation measurements from TESS. Given the $\sim$100\,Myr
decoherence time of young clusters, \replaced{the youngest ``field stars''}{these young stars} may
have easily identifiable young neighbors.

@@

Tables~\ref{tab:planets} and~\ref{tab:stars} summarize the results for
planets and stars, respectively.  Our recovered ages are accurate for
all 14 known Kepler planets in clusters.  While lithium provided
minimal added information for most of the sample\added{ due to its astrophysical
depletion rate}, $t_{\rm Li}$ and
$t_{\rm gyro}$ agreed in over 90\% of cases for which comparison
was possible.  Our \replaced{resulting ages}{results} included two-sided $t_{\rm gyro}$
and $t_{\rm Li}$ for \ltonegyrhighqconfirmedtwosided\ sub-gigayear
confirmed planets, and two-sided $t_{\rm gyro}$ with one-sided $t_{\rm
Li}$ for \ltonegyrhighqconfirmedonesided\ \replaced{sub-gigayear confirmed}{such}
planets.  Allowing for ``candidate'' planets, grazing transit
geometries, and stars with RUWE far from unity expands the counts by a
factor of two.
